\textbf{ Spectrophotometric parallax:\\ Accurate estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-08-01 paper/sp.tex
title tweaks; notes to self

% To-do items
% -----------
% - draft introduction
% - draft discussion
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - address the non-convexity of the model (see twitter convo with @andy_l_jones).
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.

@@

\shorttitle{spectrophotometric distance estimates}
\shortauthors{hogg, eilers, rix}
%\watermark{kittens}

@@

\title{\textbf{%
Accurate spectrophotometric estimates of luminosity and distance\\
for luminous red-giant stars%
}}

@@

\begin{abstract}\noindent
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which geometric parallaxes are not precisely measured by \gaia.
Here we use the \apogee--\gaia\ overlap to train a purely linear model of
continuum-normalized \apogee\ spectroscopy
(plus \gaia, \zmass, and \wise\ photometry)
that predicts absolute magnitude (logarithmic luminosity) for red stars with $0<\logg<2.2$.
The training makes use of a justified likelihood function for the \gaia\ parallaxes,
so that the training can be performed with full use of uncertainties and
without any cuts on parallax or parallax signal-to-noise ratio.
There are no physical inputs to the model; it is a pure linear regression designed
to accurately predict astrometric parallaxes.
The model includes an L1 regularization that zeros out the contributions of
most spectral pixels to the parallax estimates.
The training is performed
with leave-out subsamples such that no star's astrometry is used even indirectly in its
spectroscopic parallax estimate.
The model is flexible enough---and has enough photometric information---to
correct for extinction by dust without any
input of explicit extinction or reddening estimates.
We re-label each star in the sample
with a new spectroscopic parallax; this new parallax has an empirical (1-sigma)
uncertainty of 9~percent, which is more precise than the \gaia\ parallax
for the vast majority of targets.
Validation with globular and open clusters shows that the spectroscopic parallaxes
are both accurate and precise.
These distance and luminosity estimates open up great opportunities for
mapping the Milky Way disk with the next
generation of spectroscopic surveys, and especially \sdssv.
\end{abstract}

@@

@@

The linear model we use has many limitations: Linear models are inflexible.
However, they have a great advantage, which is that it is easy to look inside
the model and check whether the dependencies represented there make sense.
One nice thing about the model is that the feature vector $x_n$ for star $n$
contains a set of photometry in magnitudes.
Linear combinations of these photometric measurements are like complex synthetic
magnitudes or colors.
Since the model was trained to deliver luminosities of stars given training data
of different metallicities and extinctions, the synthetic photometry developed
by the model should be, in some sense, metallicity- and extinction-independent
photometry for luminous red-giant stars!
HOGG: CONTINUE THAT THOUGHT.

HOGG: Brief comment on the flexibility of the photometric model and interpretation of the photometric coefficients. Maybe plot some photometric properties vs metallicity or dust?

>>>>

% To-do items
% -----------
% - draft introduction
% - draft discussion
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - address the non-convexity of the model (see twitter convo with @andy_l_jones).
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - related: Switch the model to be a model for parallax directly.
% - does the model need a cute name, like The Swan?

@@

\shorttitle{spectrophotometric parallax}
\shortauthors{hogg, eilers, rix}
%\watermark{kittens}

@@

\title{\textbf{%
Spectrophotometric parallax:\\
Accurate estimates of luminosity and distance\\
for luminous red-giant stars
}}

@@

\begin{abstract}\noindent
With contemporary infrared spectroscopic surveys like \apogee,
red-giant stars can be observed to distances and extinctions
at which geometric parallaxes are not precisely measured by \gaia.
Here we use the \apogee--\gaia\ overlap to train a purely linear model of
continuum-normalized \apogee\ spectroscopy
(plus \gaia, \zmass, and \wise\ photometry)
that predicts distance modulus (logarithmic parallax or distance)
for red stars with $0<\logg<2.2$.
The training makes use of a justified likelihood function for the \gaia\ parallaxes,
so that the training can be performed with full use of uncertainties and
without any cuts on parallax or parallax signal-to-noise ratio (not even any non-negative
requirements).
There are no physical inputs to the model; it is a pure regression designed
to accurately predict astrometric parallaxes.
The model includes an L1 regularization that zeros out the contributions of
most spectral pixels to the parallax estimates.
The training is performed
with leave-out subsamples such that no star's astrometry is used even indirectly in its
spectrophotometric parallax estimate.
The model is flexible enough---and has enough photometric information---to
correct for extinction by dust without any
input of explicit extinction or reddening estimates.
We re-label each star in the sample
with a new spectrophotometric parallax; this new parallax has an empirical (1-sigma)
uncertainty of 9~percent, which is more precise than the \gaia\ parallax
for the vast majority of targets.
Validation with globular and open clusters shows that the spectroscopic parallaxes
are both accurate and precise.
These distance estimates open up great opportunities for
mapping the Milky Way disk with the next
generation of spectroscopic surveys, and especially \sdssv.
\end{abstract}

@@

\section{Why and how should we model red-giant stars?}

HOGG: We need to, for \sdssv.

HOGG: We know we can because of physical arguments.

HOGG: What kinds of inputs do we need to predict parallaxes?

@@

The linear model we use has a big limitation: Linear models are inflexible.
However, it has a great advantage over more general methods,
which is that it is easy to look inside
the linear model and check whether the dependencies represented there make sense.
One nice thing about the model is that the feature vector $x_n$ for star $n$
contains a set of photometry in magnitudes.
Linear combinations of these photometric measurements are like complex synthetic
magnitudes or colors.
Since the model was trained to deliver luminosities of stars given training data
of different metallicities and extinctions, the synthetic photometry developed
by the model should be, in some sense, metallicity- and extinction-independent
photometry for luminous red-giant stars!
In detail this synthetic magnitude is something like HOGG PUT EQUATION HERE AND
SAY SOMETHING ABOUT IT. HOGG: Maybe plot this synthetic photometry against dust
and metallicity.
